---
title: "Word Prediction - v1.0"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(leaflet)
library(ggplot2)
library(mgcv)
library(wordcloud)

# Load dataset
df <- readRDS("./data/3gram_en.rds")

```

Inputs {.sidebar}
=======================================================================

```{r}

textInput("words", label="Type your text here (only english):")
actionButton("go",label = "Get Next Word", icon = NULL, width = '150px')


```

Next Word
=====================================

```{r}

nextWord <- eventReactive(input$go,{
  
  #get 2 last grams and lower
  str <- input$words
  str <- tolower(str)
  str <- strsplit(str, " ")
  a <- paste(str[[1]][length(str[[1]])-1],str[[1]][length(str[[1]])])
  s <- paste("^",a,sep="")
  
  nextW <- head(df[grep(s, df[,1]),], 15)
  
  return(nextW)
  
})


# using shiny reactive
next1 <- reactive({
  
 
  ret <- nextWord()[1,1]
  freq <- nextWord()[1,2]
  str <- strsplit(ret, " ")
  str[[1]][length(str[[1]])]
  

})

next2 <- reactive({
  
  ret <- nextWord()[2,1]
  freq <- nextWord()[2,2]
  str <- strsplit(ret, " ")
  str[[1]][length(str[[1]])]
  
})

next3 <- reactive({
  
  ret <- nextWord()[3,1]
  freq <- nextWord()[3,2]
  str <- strsplit(ret, " ")
  str[[1]][length(str[[1]])]
  
})

next4 <- reactive({
  
 ret <- nextWord()[4,1]
 freq <- nextWord()[4,2]
 str <- strsplit(ret, " ")
 str[[1]][length(str[[1]])]
  
})


```

Row
-------------------------------------

### Next Word 1st {.value-box}

```{r}
# Emit the download rate
renderValueBox({
  valueBox(
    value = paste(next1()),
    icon = "fa-text-height",
    color = "primary"
  )
})

```


### Next Word 2nd {.value-box}

```{r}
# Emit the download rate
renderValueBox({
  valueBox(
    value = paste(next2()),
    icon = "fa-text-height",
    color = "green"
  )
})

```


### Next Word 3rd {.value-box}

```{r}
# Emit the download rate
renderValueBox({
  valueBox(
    value = paste(next3()),
    icon = "fa-text-height",
    color = "Orange"
  )
})

```


### Next Word 4th {.value-box}

```{r}
# Emit the download rate
renderValueBox({
  valueBox(
    value = paste(next4()),
    icon = "fa-text-height",
    color = "brown"
  )
})

```



Row
------------------------------------
   
### Word Map

```{r}


renderPlot({
  
  df <- nextWord()
   

  wordcloud(df$ngrams, # words
          df$freq, # frequencies
          scale = c(4,1), # size of largest and smallest words
          min.freq=10,
          colors = brewer.pal(8,"Dark2"), # number of colors, palette
          rot.per = 0.5) # proportion of words to rotate 90 degrees
  
  
})


```  


Help
=====================================

### Help

#### How to use?
- Select State or All to view all Country.
- Select Ranges to view results in Dashboard.
- Circles represents the amouth of population for each city.
- Click in circles to view some detais about city.


#### General Info
- Developed by: **Expedito Pinto de Paula Junior**
- Course: **Capstone**
- Description: This Data Product it was developed using *flexdashboard* and *shiny*
- Dataset Source: The dataset used from [kaggle -  brazilian cities](https://www.kaggle.com/crisparada/brazilian-cities)
- Version: 1.1
- Date: Nov 20, 2020